---
title: "case study in getting data from BOM"
author: "Nicholas Tierney"
date: "7 March 2016"
output: pdf_document
---

These are the steps I go through to get data from BOM

- I go to the website: http://www.bom.gov.au/climate/data/?ref=ftr
- "select using text"
- Select Rainfall
- Select Daily Observations
- Select a weather station in the area of interest: "Brisbane"
- Choose the matching towns: "Brisbane, QLD, 27.47ºS, 153.03ºE"
- Choose Nearest Bureau stations: "040245 Toowong Bowls Club QLD (4.4km away)
- "Get data"
- select "All years of data"
- Open ZIP file
- read into R

```{r knitr-setup, include = F}

knitr::opts_knit$set(root.dir = '..')

```

Then I read in the data, get a sense of what the data looks like

```{r message = F, warning = F}

library(readr)

dat_toowong <- read_csv("IDCjAC0009_040245_1800_Data.csv")

dim(dat_toowong)

library(dplyr)

glimpse(dat_toowong)


```

Then collect the date information into a format that R understands

```{r message = F, warning = F}

# combine year-month-day into date string

library(lubridate)

dat_clean <- 
  dat_toowong %>%
  mutate(date_time = paste(Year, Month, Day, sep = "-")) %>%
  mutate(date_time = ymd(date_time, tz = "Australia/Brisbane")) %>%
  rename(rainfall_in_mm = `Rainfall amount (millimetres)`)

```

then plot this in R


```{r message = F, warning = F}

library(ggplot2)

ggplot(data = dat_clean,
       aes(x = date_time,
           y = rainfall_in_mm)) + 
  geom_line() + 
  labs(title = "rainfall over time in the Toowong Bowls Club region")

```

The good things are that the data from BOM is in a nice format, and required very little adjustment to get a nice plot.

However I think it would be nice to automate this from within R, so you could instead write something (roughly) like below:

```{r eval = FALSE}

library(bomr)

bomr_get_locations(location = "brisbane",
                   type = "rainfall",
                   frequency = "daily")

# list of locations around Brisbane.
# 

data <- 
bomr_get_data(long = "27.47S",
              lat = "153.03E",
              type = "rainfall",
              frequency = "daily",
              date_start = "01/01/2016",
              date_end = "01/29/2016")

```

It could also be possible to merge your own dataset with date and long/lat info with BOM data, for example:

```{r eval = F}

new_data <-  bomr_merge_data(new_data = my_data,
                             type = "rainfall", 
                             frequency = "daily")

```

Which would avoid the user needing to do their own date matching on their own computer.
